---
title: 'Berliner Hotels Don`t Have To Fear Airbnb - At Least For Now'
subtitle: "Paulo Kalkhake & Dan Murphy Hertie School of Governance"
output: 
  html_document:
    toc: true
    linestretch: 2
    toc_float:
      smooth_scroll: TRUE
    highlight: monochrome
bibliography: Bibliography.bib

---

```{r chunk_options, include=FALSE, results='asis'}
knitr::opts_chunk$set(cache=TRUE, fig.path = "images/")

try(setwd("/Users/Paulo/GitHub/FinalPaper/website"),silent=TRUE)
try(setwd("/Users/djm113/Documents/GitHub/FinalPaper/website"),silent=TRUE)

library(googleVis)
suppressPackageStartupMessages(library(googleVis))

```


``` {r, include=FALSE, results='asis'}
source("/Users/Paulo/GitHub/FinalPaper/code/finalpaper_analysis.R")
```

<br>
<br>
***Airbnb claims to be part of the sharing and disrupting the hotel industry. However, data shows that the majority of Airbnb listings in most cities are entire homes, many of which are rented all year round - disrupting housing and communities. Browse the data for Berlin below, and see for yourself.***
<br>
<br>

Airbnb and other short-term rental platforms for apartments have become hugely popular - particularly among yonger traveler, offering zaccommodation that is cheaper, better located or more interesting than hotels. With its growing success, the effects of Airbnb on the Berlin housing market has recently been the subject of a great deal of scrutiny. 


Airbnb claims that their service is complementary to hotels. However, data shows that the majority of Airbnb listings in most cities are entire homes, many of which are rented all year round - disrupting housing and communities. With that debate in mind we looked at what data might tell us. Browse our results below, and see for yourself.

## Airbnb has changed the accomodation industry forever - and keeps growing

In 2014, the city passed a law that forbade renting out entire apartments to tourists on Airbnb. This came in response to an unprecedented 56 percent rise in rents in Berlin between 2009 and 2014, for which Airbnb and other online hosting platforms received some of the blame. The city's law was widely welcomed by Berliners hoping to slow the forces of gentrification, but it was also welcomed by another Airbnb victim in Berlin; the hotel industry. 

```{r, echo=FALSE, include=TRUE}

GeoBerlin_bev <-gvisGeoChart(analysis_simple,
                locationvar = 'neighbourhood',
                colorvar = 'AB_supply',
                options = list(region = "DE", 
                              displayMode="regions",
                              resolution = 'provinces',
                              showLegend = TRUE,
                              width=600, height=400))

plot(GeoBerlin_bev)

```


## Airbnb Listings per Neighbourhood (static model) (2010 - 2014)
```{r, echo=FALSE, fig.height=4.5, fig.width=8}


```

## How big is Airbnb in Berlin?

While Berliners fought tourists for apartments, hotels were quietly fighting Airbnb for tourists. That's a real challenge for the long-time incumbents, whose average price of 80 euros per night cannot compete with the average Airbnb price of 55 euros per night. 


To make matters worse, the rise of Berlin as the place to be happened to coincide with the rise of Airbnb as the place to stay. This prevented the hotels of Berlin from fully realizing profits that were, in their minds, rightfully theirs. 

## Does the number of Airbnb Apartments have an effect on Hotel occupancy in Berlin?



****Insert Cool Map of Berlin here****

### But remember, Berlin's popularity increased over the same period
****Insert Descriptive graph of popularity control****

## So, is Airbnb hurting hotels in Berlin?

### Effect of Increase in Airbnb Listings on Berlin Hotel Occupancy Rates (2010 - 2014)
```{r, echo=FALSE, fig.height=3.8, fig.width=8}

```

### Estimation Results {.flexbox .vcenter .reduced}
``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}


```

### Estimation Results (2) ## {.flexbox .vcenter .reduced}
``` {r, echo = FALSE, include=TRUE, message = FALSE, error = FALSE, results = "asis"}


```


## What else is Airbnb doing to our cities?
Our statistical analysis  






## Useful table format

                    not reelected   reelected
------------------ --------------  ----------
  non-board member      182 (17%)   866 (83%)
      board-member       41 (14%)   252 (86%)

##Questions? 



## Usefule References Format

[^1]: The presented research project is part of a paper project by Guillermo Rosas (Washington University in St. Louis) and Jonas Markgraf (Hertie School of Governance Berlin).

[^2]: Spanish public savings banks (*cajas*) faltered and dissolved in the course of the bursting Spanish real estate bubble ([link](http://voxeu.org/article/five-lessons-spanish-cajas-debacle-new-euro-wide-supervisor)); similarly, German *Landesbanken* got into serious trouble and were partly consolidated ([link](http://www.sueddeutsche.de/geld/fehlspekulationen-der-landesbanken-chronologie-einer-krise-1.283250) (in German)). At the same time, German *Sparkassen* weathered the financial crisis well and emerged strengthened from it ([link](http://www.telegraph.co.uk/finance/financialcrisis/3190674/Financial-crisis-boosts-German-banks.html)). This heterogeneous crisis performance of public banks remains largely unexplained and motivates the PhD project of one of the authors.